---
title: "Project Group 1"
author: "Adil Iqbal, Angelita Krepel"
date: "11/22/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1

### Introduction

This data was extracted from the U.S. Census in 1994. Our goal will be to predict whether an individual earns more than $50,000 USD per year (`class`) based on data collected by the census alone. Examples of data collected by the census include education level (`education`), marital status (`marital.status`), and country of origin (`native.country`) among others. All features are either categorical or continuous integer values.


#### Features

age: continuous.
workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.
fnlwgt: continuous.
education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.
education.num: continuous.
marital.status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handl sdasdasders-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.
relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.
race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.
sex: Female, Male.
capital.gain: continuous.
capital.loss: continuous.
hours.per.week: continuous.
native.country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands.

#### Response Variable

class: >50K, <=50K.

#### Question
Can we predict if an individual will make 50K or more based on a list of predictors?

```{r}
set.seed(1)
data = read.csv("adult.csv")

summary(data)
```

### Methods

#### Removing missing values. 

In this data set, missing values are marked with a "?" character. Since all the missing data occurs in our categorical variables, it does not make sense to replace them with a median or a mean. So we are opting to remove them entirely for our analysis. After removing the observations with missing variables, we are left with 45,222 observations.

```{r}
#Adil
cols_missing_vals <- c("occupation", "workclass", "native.country")

for (column in cols_missing_vals) {
  data <- data[!grepl("\\?", data[, column]),]
}

summary(data)
```

#### Set categorical features to dummy variables
In this data, many categorical variables are needed to be converted to numeric and make a Dummy variable for the response variable.

```{r, results='hide'}
#Adil & Angelita 
### 1. Class  (<=50K or 50K>)
data$class = ifelse(data$class==" <=50K",0,1) # Our Response Variable. I made 0 for below 50K and 1 for above 50K


### 2. Sex  (Female, Male)
data$sex = ifelse(data$sex==" Female",0,1)



### 3. Workclass
data$workclass = ifelse(data$workclass==" Private",0,
                                  ifelse(data$workclass==" State-gov",1,
                                         ifelse(data$workclass ==" Federal-gov",2,
                                                ifelse(data$workclass == " Self-emp-not-inc",3, 
                                                       ifelse(data$workclass==" Self-emp-inc",4, 
                                                              ifelse(data$workclass==" Local-gov",5,6)))))) #



### 4. Marital Status
data$marital.status = ifelse(data$marital.status == " Widowed",0,
                                   ifelse(data$marital.status == " Divorced",1,
                                          ifelse(data$marital.status == " Separated",2,
                                                 ifelse(data$marital.status == " Never-married",3,
                                                        ifelse(data$marital.status== " Married-civ-spouse",4,
                                                               ifelse(data$marital.status==" Married-AF-spouse",4,5
                                                                      ))))))




### 5. Race
data$race = ifelse(data$race ==" White",0,
                         ifelse(data$race == " Black",1,
                                ifelse(data$race==" Asian-Pac_islander",2,
                                       ifelse(data$race==" Amer-Indian-Eskimo",3,4))))
### 6. Relationship
data$relationship = ifelse(data$relationship==" Husband",1,
                                 ifelse(data$relationship==" Wife",1,0))



### 7. Education
# I grouped them by following:

# 0 means all students who are in 1st to 12 grade
# 1 means all students who are in HS-grad, Assos-acdm, Prof-school
# 2 means all students who are in Bachelors Degree
# 3 means all students who are in Masters' Degree
# 4 means all students who are in Doc 
# and 5 means UNKNOWNs
data$education=ifelse(data$education==" Prepschool",0,
                            ifelse(data$education==" 1st-4th",0,
                                   ifelse(data$education== "5th-6th",0,
                                          ifelse(data$education==" 7th-8th",0,
                                                 ifelse(data$education==" 9th",0, ifelse(data$education==" 10th",0,
                                                                                                ifelse(data$education==" 11th",0,
                                                                                                       ifelse(data$education==" 12th",0,
                                                                                                              ifelse(data$education==" HS-grad",1,
                                                                                                                     ifelse(data$education==" Assos-voc", 1,
                                                                                                                            ifelse(data$education==" Assoc-acdm",1,
                                                                                                                                   ifelse(data$education==" Prof-school",1,
                                                                                                                                          ifelse(data$education==" Some-college",1,
                                                                                                                                                 ifelse(data$education==" Bachelors",2,
                                                                                                                                                        ifelse(data$education==" Masters",3,
                                                                                                                                                               ifelse(data$education==" Doctorate",4, 5))))))))))))))))

#print(data$education_dummy)


### 8. Native Country

data$native.country = ifelse(data$native.country==" United-States",0,1)
# So most people are from the United States, so I assigned them to 0 and other countries to 1


# Omit NA rows
data = na.omit(data)
```

```{r}
#Angelita
#turning occupation into dummy variables
#install.packages('fastDummies')
library('fastDummies')
data <- dummy_cols(data, select_columns = 'occupation')
```
```{r}
#drop original occupation column to keep new dummies
data <- data[-c(7)]
```


#### Split data into training and testing set

Below, we are now splitting the data into training and testing sets. With the training set containing 80% of the data and the testing set containing 20%. 

```{r}
#Dosbol
sample <- sample(c(TRUE, FALSE), nrow(data), replace=TRUE, prob=c(0.8,0.2))
train  <- data[sample, ]
test   <- data[!sample, ]

print(nrow(train))
print(nrow(test))
```


```{r}
#checking correlations
pairs(adult[sapply(adult, is.numeric)])
cor(adult[sapply(adult, is.numeric)])
```
Doesn't appear that from the original model there is much correlation between the data, but after making the occupation dummies there may be some correlation between those.


#### Model 1: Logistic regression

We will be using logistic regression, because the response variable is a categorical variable that takes 1 if their salary is >=50K.

```{r, warning=FALSE}
#Angelita
lr.glm = glm(class ~., data = train, family = "binomial")
glm.pred = predict.glm(lr.glm, test, type = "response")
yHat = glm.pred > 0.5
table(test$class,yHat)
```
```{r}
(1485+1640)/(3078+1485+1640+2807)
```
The test error rate is very large sitting at about 35%. This means our model only predicted about 65% of the data correctly. 

```{r}
#Angelita
summary(lr.glm)
```
Based on the summary of the logistic regression model we can see thatnot all the variables are important in predicting, so we will try again but this time with the varaibles deemed important.

```{r}
#Angelita 
lr.glm2 = glm(class ~  age+fnlwgt+education.num+marital.status+relationship+sex+capital.gain+capital.loss+hours.per.week+`occupation_ Adm-clerical`+`occupation_ Exec-managerial`+`occupation_ Farming-fishing`+`occupation_ Other-service`+`occupation_ Priv-house-serv`+`occupation_ Prof-specialty`+`occupation_ Protective-serv`+`occupation_ Sales`+`occupation_ Tech-support`, data = train, family = "binomial")
glm.pred2 = predict.glm(lr.glm2, test, type = "response")
yHat2 = glm.pred2 > 0.5
table(test$class,yHat2)
summary(lr.glm2)
```
```{r}
(1484+1627)/(3079+1484+1627+2820)
```
This did about the same. 
Creating a decision tree model to predict the class variable


####Decision Tree Model
```{r}
library(tree)
plot.new()
class.tree = tree(class_dummy ~ native.country_dummy+education_dummy+relationship_dummy+race_dummy+marital.status_dummy+workclass_dummy+sex_dummy+capital.loss+capital.gain+age, data=train)
summary(class.tree)
plot(class.tree)
text(class.tree)
```
